\chapter{Judges and Officials Rules}

\section{Street Director}

The Street Director is the head organizer and administrator of Street.
With the convention host, they are responsible for the event logistics, the equipment and the system used to run the event.
They must select the Chief Judge.
They are in charge of keeping the event on schedule and answering questions about the event.
The Street Director is the highest authority on everything to do with the Street competition, except for decisions on rules and results.

\section{Chief Judge}

The Chief Judge is the head Street official.
They are responsible for selecting the judges, overseeing the competition, handling protests, ensuring the event rules are followed, and answering questions about the rules and judging.
The Chief Judge is also responsible for the accuracy of all judging point tabulations and calculations.

An interruption of judging can result from material damage, injury of a competitor, or interference with a competitor by a person or object.
The Chief Judge determines if the rider is at fault.
If they are not, the Chief Judge chooses when to insert the rider back into the competition, along with the rider's remaining time.
The Chief Judge may be the same person as the Street director.

\section{Judges}

\subsection{Judging Panel}

There are three judges per section for the preliminary rounds, and five judges for the finals.

\subsection{Selecting Judges}
A person should not judge an event if he or she is:
\begin{itemize}
\item A parent, child or sibling of a rider competing in the event.
\item A coach, manager, trainer or colleague of the same club/team as a rider competing in the event.
\item More than one judge from the same on the same judging panel.
\end{itemize}

If the judging pool is too limited by the above criteria, restrictions can be eliminated starting from the bottom of the list and working upward as necessary, but only until enough judges are available.

\subsection{Judging Panel May Not Change}

The individual members of the judging panel must remain the same for an entire category.
In the event of an emergency, this rule can be waived by the Chief Judge.

\subsubsection{Rating Judge Performance}
Judges are rated by comparing their scores to those of other judges at previous competitions.
If a judge's performance is determined to be too weak, they may be removed from the judging panel.

Characteristics of Judging Weaknesses:
\begin{itemize}
\item \textbf{Excessive Ties:}
Using ties frequently (it defeats the purpose of judging.)
\item \textbf{Bias:}
Placing members of certain groups or nations significantly lower or higher than other judges.
\item\textbf{Inconsistence:}
Ranking a large number of riders significantly different from the average of other judges.
\end{itemize}

\subsection{Training}
The judges workshop is set by the Street Director or Chief Judge.
Either the Street Director or the Chief Judge run the workshop.
The workshop must be help before the competition.
Judges should have read the rules prior to the start of the workshop.
The workshop will include a practice session.
Each judge will read the rules, attend the workshop, agree to follow the rules and agree to their potential removal from the list of available judges if they show excessive judges weaknesses, as determined by the Chief Judge.

\subsection{Street Comp Judging}
There will be three judges for each zone for the preliminaries, and five judges for each zone for the final.

The judges will have 5 minutes after each session to discuss the tricks.
They should use their preferred system for taking notes during competition (for example: they may take notes of tricks that were landed in that zone or assign point values).
It is recommended to ask riders about specific tricks (other riders should be present to justify the response), but neither the riders nor judges may discuss relative difficulty as it could influence the score.
After the judges have seen ALL of the riders at a single zone, they will rank the riders from best to worst.
Each judge is responsible for one set of rankings at their zone; they must judge all riders against each other even if they are in different groups.

For prelims, the riders will be given points according to their placement.
(So 1st place gets one point, 2nd place gets two points, etc.)
For finals, each place is awarded points as follows:

\begin{tabular}{|l|l|}
\hline
\textbf{1st}& 10 points\\
\hline
\textbf{2nd}& 7 points\\
\hline
\textbf{3rd}& 5 points\\
\hline
\textbf{4th}& 3 points\\
\hline
\textbf{5th}& 2 points\\
\hline
\textbf{6th}& 1 point\\
\hline
\textbf{7th and Beyond}& 0 points \\
\hline
\end{tabular}

The ranking should be influenced by the number of tricks done, and the difficulty of the tricks.
Consistency should not be considered, because it is inevitable that a consistent rider will land the most tricks.
However, note that the number of tricks
should also not always be the deciding factor on who wins.
Some one who performs 18 easy tricks should not be scored higher than someone who performs 3 outstanding tricks.
Once the judges assign places for every zone, the points will be added up and the results can be calculated.
After prelims, the riders with the lowest sum of placement points move on to finals.
In finals, the rider with the most points is the winner.

\subsection{Sportsmanship}
If a rider distracts or delays the competition or shows unsportsmanlike conduct, the Chief Judge may choose to warn or eliminate that rider.
