\chapter{Event Organizer Rules}

\section{Venue}

\subsection{Street Comp Performing Area \label{sec:flat-street_street-performing-area}}
The Street course is to be composed of three ``zones''.
Each zone should have multiple obstacles, but each obstacle should encourage a specific type of skill.
The list below is an example of three typical things that can be used for the zones; however designers of the Street comp area should not limit themselves to the exact list.

\textbf{Zone 1:}
A ramp with a skate park rail in the middle, and a ledge on either side.
This zone will encourage technical grinds, without giving an advantage to a right or left footed grinder.

\textbf{Zone 2:}
Two different manny pads (a smooth platform of at least 3 m x 0.5 m and between 7 cm and 15 cm in height), one with two revs of length, and one with just one rev of length.
This will encourage the ability to perform technical flip tricks and other Street moves while having to set up quickly for the move down.

\textbf{Zone 3:}
A set of 5 stairs and a set of 7 stairs with a handrail in the middle of each (that are of a similar size to one that you would find in a city, not extremely steep).
This section would encourage the ability to perform bigger moves of all types.

Depending on the time limit of a host and physical limitations, elements described above can also be used in one bigger zone to give a rider more flexibility.
A host should be aware then that this needs 3 times as much time for the competition and the host should also be sure that judges are able to follow the rider in the entire big zone.
It is also possible to use a real street environment if that is possible.
This may result in having some different obstacles than specified above but it provides a 100\% real street atmosphere.

Independent from the setup a host can go for, they should always take care to offer room for technical street, for grinds and for some big stuff.
He should also pay attention to offer enough room for the big wheelers to go for Big Street.
The descriptions of the zones above should give a good idea about the requirements while offering a lot of room for being creative.

\subsection{Problems With Required Obstacles}
The required obstacles must be built strong enough to endure many hours of heavy use.
They need to survive the competition without changing their shape or stability.
If one of the required obstacles is broken or made unusable during the competition, it must be repaired if one or more competitors say they need to use the damaged part.
If no competitors have a problem with the damage, no repair is necessary except for safety reasons, such as in the event of sharp exposed parts.

\subsection{Postponement due to Weather}

In the case of rain or bad weather and an uncovered Street area, the organizers should postpone the event and exchange all the affected parts of the course for dry ones (replacing pallets for example).
The competition should be canceled if it is considered dangerous for the riders.
If the event is postponed or moved to an indoor location, allowances must be the same (metal pedals, marking tires, etc.)
The event host should try to place events that may be influenced by weather conditions in the first days of the event, giving a larger period of time to reschedule it.

\subsection{Music}
In Flatland, a DJ plays music for the competition.

\section{Officials}

The host must designate the following officials for street:
\begin{itemize}
\item Street Director
\item Chief Judge
\end{itemize}

The host must designate the Street Director well in advance of the event.
For an international events, it is recommended that the Street Director is chosen at least one year in advance so that they may be consulted on scheduling.
The Street Director must select the Chief Judge.
The Chief Judge may be the same person as the Street director.

\section{Communication}

Hosts must publicize details of the available competition area as far in advance of the competition as possible, and organizers of international championships at least three months prior to the event.
For other events, the organizers must specify the venue for the Street competition by the beginning of the convention/competition at the latest.

\section{Categories}
There are no mimimum categories other than male and female.
If there are less than three females or less than three males overall, the male and female categories are merged.

\section{Practice}

Event organizers must arrange that the course for the Street competition is set up and available to be practiced on before competition.
With different time frames depending on the time frame and duration of the convention/competition.
At least 2 days prior to the day of the competition should apply to events that are over a weekend long (4 days and over of competition).
For events that last 1 - 3 days, practice time can be at the discretion of the competition organizers, for example, prior, but on the day of the competition, practice time must still be allowed.
If practicing on the competition grounds is not possible prior to the competition day, the organizers must build similar objects on another location for the riders to train on.
