\chapter{Overview}

\begin{framed}
New text, not from the 2014 rulebook is in \textbf{boldface}. Text that we suggest be deleted is in \textit{italic}. Since the text has been reorganized, the original 2014 sections are marked with a boxed reference, such as \oldrule{3.4}. Meta-text, intended for the rulebook editors is boxed.
\end{framed}

\section{Definition}

\textbf{Flatland is a skills competition where riders perform 
tricks on a flat surface (usually pavement). Judging is based solely on
technical skill, including combinations and transitions between skills.
The competition consists of preliminary rounds, followed by a battle-style 
final.}

\label{sec:flat-street_flatland-overview}}

\oldrule{6.13}
\textit{A Flatland skill is any unicycle skill performed on a flat surface.
Flatland encourages riders to demonstrate a high level of technical difficulty and variety, as well as combinations and transitions between skills.}

\oldrule{6.1}
\textit{In Flatland, riders perform in a flat area with no obstacles or props.
There is no judging of music and costume, and the emphasis is on originality and creativity.}

\section{Rider Summary}

\textbf{This section is intended as an overview of the rules, but does not
substitute for the actual rules.
\begin{itemize}
\item Be aware of the deadline for entry.
\item You must be on time for your scheduled performance.
\item You must wear shoes. No additional safety equipment is needed.
\item You may bring your own music, but music is not judged.
\item Preliminary round performances are for one minute.
\item Battles are for two minutes.
\item Final round battle duration is determined by the host, but is usually three minutes.
\end{itemize}}