\chapter{Event Organizer Rules}

\section{Venue}

In the Speed Trials Comp, the organizers should postpone the event in case of wet weather and exchange all the affected parts of the course for dry ones (replacing pallets for example).
These competitions should be canceled if considered dangerous for the riders.
If postponed or moved to an indoor location the organizers must try to keep the allowances the same as outdoors competitions (metal pedals allowed for example).
If originally on the competition schedule, these canceled competitions should be rescheduled during the convention duration.
The event host should try to place events that may be influenced by weather conditions in the first days of the event, giving a larger period of time to reschedule it.

There should be no dangerous objects to land on if a rider falls off a high object.
Sections should be constructed so that they do not collapse or fall over under normal riding conditions.

\section{Officials}

The host must designate the following officials for Speed Trials:
\begin{itemize}
\item Speed Trials Director
\item Chief Judge
\end{itemize}
Starter, Line Judges and Timekeepers/Finishers will be volunteers assigned during the event.
\section{Communication}

\section{Age Groups}

Male and female competitions should be offered with age group winners set as 0-9 (Kids), 10-15 (Junior), 16-34 (Adult) and 35+ (Senior).

\section{Course Design}

Construction of the course is at the discretion of the Speed Trials director.
However, the course should have a combination of long jumps, high jumps, balance elements, and precision jumps.
The finals course should be more difficult than the preliminary course and all courses must not favour a particular hopping stance or direction.
The Speed Trials director needs to provide a warm-up area with obstacles for riders.

\section{Course Planning}

Adequate time and resources is required to set up a proper course with at least one day required to build identical lines for multiple riders to complete preliminary runs concurrently.
