\chapter{Competitor Rules}

\section{Safety}
Riders must wear shoes, knee pads, gloves/wrist-guards and a helmet (see definitions in chapter \ref{chap:general_definitions}).

\section{Unicycles and Categories}

The host or event director may establish categories based upon ability.
The categories should be named accordingly, if used: Beginners, Intermediate, Elite.
Elite is the only category which decides the overall champion.

Categories should not separate competitors by any kind of unicycle setup (wheel size, crank size, gearing, etc.).

\section{Rider Identification}

Riders must wear their race number(s) fixed at the four corners, clearly visible on their chest and, when required, on their back so that it is visible during the race and as the rider crosses the finish line.
Riders must use the officially provided race number unmodified in any manner.
Numbers should not be folded, trimmed, or otherwise defaced.
Referee approval must be sought to modify a number plate if it cannot otherwise be attached securely due to hydration pack, rider physique or posture when riding.
Lost or damaged race numbers must be replaced with approval by referee.
Riders who lose or damage their race numbers \textit{during} a race will not be penalized.
Additionally, riders may be required to wear a chip for electronic timing.

\section{Protests}

Protests must be filed on an official form within two hours of the posting of event results.
Every effort will be made for all protests to be handled within 30 minutes from the time they are received.

\section{Event Flow}

\subsection{Starting}

There will be a Le Mans style start.
Unicycles will be lined up in a designated area away from the riders near the lap/finish line.
Riders will line up behind an additional line and then be required to run to retrieve their unicycle when the race starts.
They will then need to mount their unicycle to ride.
Riders must be mounted within 10 meters after crossing the lap/finish line.

\subsection{Passing}

An approaching rider needs to alert the slower rider of their intentions to pass.
Special care at international events should be taken due to language differences.

\subsection{Dismounts}

Upon dismounting there are no restrictions about passing riders.
Dismounted riders may run with their unicycle.
Courtesy is expected to avoid accidents, but the running unicyclist does not have to yield to riding unicyclist.

\subsection{Illegal Riding}

Riders cannot cut the course around the obstacles.
They may ride through the obstacle section if possible or dismount and run with their unicycle.
By definition, the majority of riders should not be able to ride or hop the obstacle section.
Riding or hopping through the obstacle section should not damage or break the obstacle.

Unicyclists must attempt to ride at least 50\% of the course on each lap.
This is to avoid someone running the entire race carrying or pushing a unicycle without riding it.
A racer in violation will be warned by a racing official.
Failure to heed the warning will result in a disqualification.

\subsection{Finishes}

Riders can cross the finish line mounted, as well as walking or running.
