\chapter{Judges and Officials Rules}

\section{Muni Director}

The Muni Director is the head organizer and administrator of muni events.
With the Convention Host, the Muni Director determines the course, obtains permits, interfaces with the community, and determines the system used to run the event.
The Muni Director is responsible for the logistics and equipment for all muni events.
With the Referee, the Muni Director is in charge of keeping events running on schedule, and answers all questions not pertaining to rules and judging.
The Muni Director is the highest authority on everything to do with the muni events, except for decisions on rules and results.

\section{Referee}

The Referee is the head muni official, whose primary job is to make sure the competitors follow the rules.
The Referee makes all final decisions regarding rule infractions.
The Referee is responsible for resolving protests.

\section{Starter}

The starter starts races, explains race rules, and calls riders back in the event of false starts.
The starter is in charge of checking riders for correct unicycles and safety equipment.

\section{Course Marshall}
The Course Marshall is responsible for securing and marking the course, as well as for positioning and instructing volunteers to signal the correct way to race participants or to serve as judges in the case of dismounts (see respective rule sections for details).  Furthermore, the Course Marshall should assist the Muni Director in selecting an appropriate course abiding by the guidelines in this rulebook.

\section{Officials Can Compete}

The Referee may not compete in any competition where they may be required to make a decision.
The Muni Director may compete, as long as the race course has been announced early enough that the Muni Director does not have an advantage from knowledge of the course.
