\singlechapter{Jumps \label{chap:Jumps}}

Unicycling versions of the High Jump and Long Jump.

\section{Unicycles}
Standard unicycles must be used (see definition in chapter \ref{chap:general_definitions}).
No restriction on wheel or crank size.
Metal pedals are allowed for their strength and better grip.
This may make it impossible to hold this event on a sensitive track surface.

\subsection{Broken Unicycle}
If the unicycle breaks during an attempt, a new attempt must be given to the rider.

\section{Practice Areas}
The organizer should provide a place and equipment similar to those used for the official competition so the riders can practice before making their official attempts.
The equipment should be available during the whole length of the event, and even before if the organizer decides so.

For bigger events such as Unicon, national or continental events, the organizer must provide said equipment.

\section{Safety Gear}
For Long Jump and Long Jump on Platform, riders must wear shoes, a helmet and kneepads.
In addition, for Long Jump gloves are required.

For High Jump and High Jump to Platform, riders must wear shoes and a helmet.
Shin guards are strongly recommended while gloves and kneepads are optional.

Definitions of all safety gear can be found in chapter \ref{chap:general_definitions}.

\section{Number of Attempts}
\textbf{Prelims:}
In the preliminary round, riders have 12 attempts to complete their best performance.
These attempts can be at any distance and there is no additional limit to the number at any one distance.
However each attempt must be at the same or better distance.
Thus a rider cannot attempt a longer distance, fail, and then attempt a shorter distance.
The best successfully completed attempt is the rider's result.

\textbf{Finals:}
Riders get three attempts for each distance.
Riders must successfully complete each distance before moving on to the next distance.

\section{High Jump}
The rider and unicycle jump over a bar, without knocking it down, and ride away without a dismount.
There are three parts to a successful jump: 
\begin{enumerate}
\item Riders must mount before the start line, to show they are on the unicycle and in control.
The attempt starts when the rider crosses the start line.
The rider may break off from a jumping attempt before leaving the ground, but must then start again from behind the start line.
That attempt then doesn't count.
\item Riders must jump over the bar without knocking the bar off the apparatus.
The bar can be hit as long as it does not fall.
If the bar falls before the rider crosses the finish line, it counts as an unsuccessful attempt.
\item After landing, the rider must stay in control of the unicycle until he cross the finish line without dismounting, touching a hand to the ground or any other object, or knocking down the bar or any of the high jump apparatus.
The rider starts at a low height and after each successful attempt, the height increases at set intervals.
The maximum height that was completed is recorded as the rider's result.
\end{enumerate}

\subsection{Setup}
Around the High Jump apparatus a circle with a radius of 3 meters must be marked.
This circle is start and finish line.
The rider can cross it wherever he wants.
Riders must ride or hop across the finish line for the attempt to count.
Successfully crossing the finish line is judged the same as in racing (see section \ref{sec:track-field_finishes}).
The bar must be held loosely in the jumping apparatus so it can fall or break away if the rider does not complete the desired height.
Magnetic systems are not allowed.
The bar shall have a minimum diameter of 2cm and a maximum bending of 2 cm.
(The bar may sag no more than 20 mm at its lowest point.)
The bar must be sufficiently long such that minimum distance between the two apparatuses holding the bar is 2 meters.


\section{Platform High Jump}
The object is to jump from the ground to a platform, with no pedal or crank grabs.
Riders must remain in control of the unicycle (stay mounted) for 3 seconds after landing.

The rider and unicycle jump up to a landing surface on a platform (see setup for definitions) and remain on the landing surface for a 3-second count from a judge.
The rider begins at a low height, and after each successful attempt, the height increases at set intervals.
The maximum height that was completed is recorded as his or her result.

There are three parts to a successful jump:
\begin{enumerate}
\item Riders must mount their unicycles on the ground.
A jump attempt occurs when any part of the unicycle or rider touches any part of the platform or landing surface.
The rider may break off from a jump attempt before touching the platform or landing surface.
This does not qualify as a jump attempt.
\item Riders must jump ``to rubber'' on the landing surface.
No part of the rider, or any part of the unicycle other than the tire, may touch the platform or landing surface.
\item After landing, the rider must remain mounted and on the landing surface for 3 seconds, as counted by a judge.
The rider may do any form of idling, hopping or stillstanding during the 3 seconds.
Once the judge has counted 5 seconds, the rider may return to the ground in any fashion he or she chooses.
\end{enumerate}
If a rider completes all of the requirements listed in items 1 through 3 above, the jump is deemed successful.
Otherwise, it is deemed a failed jump attempt.

\subsection{Setup}
The structure consists of two parts: a platform and a landing surface.
The top surface of the platform must be at least 120 x 120 cm in size, the maximum and recommended size is 120 cm wide by 160 cm long.
The sides of the platform must be nearly perpendicular with the ground to ensure its presence does not hamper riders.
The landing surface consists of a flat piece of wood of the same dimensions as the platform (120 x 120 cm to 120 x 160 cm) that is firmly affixed to the top of the platform.
The method of attachment must be chosen so that it does not interfere with riders during jump attempts.
The structure should be sturdily built and shimmed so that there is minimal motion when jumps are attempted.
The front of the platform must be covered by a wooden plate that extends a minimum of 60cm from the top of the platform down.
This cover must be used when riders are jumping 60cm or higher.

A 3 meter safety ring must be marked out around all sides of the platform where no persons may enter during a jump attempt.
Organizers may choose to mark out a 4 m ring where only select persons may enter during a jump attempt (e.g. photographers or judges).
A 15 m runway must also be cleared of persons for riders that will roll into the jump.

\section{Long Jump}
The rider jumps as far as possible from a jump marker, to a landing without a dismount.
The rider must then continue riding across a finish line to show control.
Riders must clear 3 markers (jump marker, landing marker and finish line) to make the jump count.
Riders may jump with the wheel going forward or sideways.
After landing, the rider must stay in control of the unicycle for the remainder of the distance from the jump marker to the finish line without dismounting, or touching a hand to the ground or any other object.
If the tire touches the jump marker before takeoff or the landing marker, it counts as a foul.
Riders may break off in a run as long as he is between start line and jump marker but if they cross or touch the jump marker, the attempt counts, including fouls.
The farthest non-fouling, successful jump is recorded.

To avoid endless competitions, the length to jump will always increase by 5cm for each round.
Once there are only 5 riders left, it's up to the riders to decide in which steps they continue.
For each age group the minimum length should be adjusted to a useful level such as 150cm for 15+ and 70cm for 0-15.
The host can adjust this depending on the level at his competition.

\subsection{Setup}
The riding area consists of a start line, a jump marker, a landing marker and a finish line beyond the jump marker.

The finishing line should be at least 4 meters from the landing marker but no more than 8 meters away.
We suggest that judges set up the finishing line 8 meters from the jump marker and move it further away if need during longer jumps.
Riders must ride or hop across the finish line for the attempt to count.
Successfully crossing the finish line is judged the same as in racing (see section \ref{sec:track-field_finishes}).
The start line must be a minimum of 25 meters in front of the jump marker to allow the riders to accelerate.
There must be an area behind the finishing line which is a minimum of 7 meters long and 2 meters wide as safety zone.
Riders may use all or part of the 25 meters between start line and jump marker.
Riders are also allowed to start from beside to be able to do accelerated side jumps.
Markers for takeoff and landing (jump marker and landing marker) must consist of a material which cannot be deformed in order to have the same conditions for all riders.
The markers must be at least 1.20 meter in width (across the runway), no more than 10 mm in height (above the runway), and no less than 5 centimeters in depth (front to back).
A Long Jump competition needs a minimum area of 40x2.5 meters.

\subsection{Judging}
The rider must clear the jump marker and the landing marker without touching them; he also has to clear the finish line to make it a valid jump.
Jump distance is measured between the outer edges of the jump and landing marker.
There has to be at least one judge (better two) to look at the markers.
For national championships and Unicons, two judges are always needed; one to observe each marker.

\section{Long Jump on Platform}
In the Long Jump on Platform competition, the rider attempts to jump as far as possible from a short pallet run-up to a landing pad without a dismount.
Following the jump, the rider must then stay mounted and in control on the unicycle for 3 seconds on the landing pallets or in the landing half-circle (described in setup below).

Riders may jump with the wheel going forward or sideways.
The rider may break off the attempt as long as they are still on the pallet run-up.
As soon as they jump in any direction from the pallet run-up, it counts as an attempt.
The farthest non-fouling, successful jump is recorded.

To avoid endless competitions, the length to jump will always increase by 5cm for each round.
Once there are only 5 riders left, the final starts and it’s up to the riders to decide in which steps they continue. 

\subsection{Setup}
The riding area consists of a starting run-up of pallets which is between 7.5 and 8.5 meters long, 1 to 1.5 meters wide and 40 to 45 cm (3 pallets) high.
A landing platform of the same height and width as the run-up is required.
The landing platform should be approximately 4.5 meters long and fixed together so that they do not move when landed on.
It is recommended to cover the pallets with plywood or a similar material.
Behind the landing platform a half circle with 3 meter in diameter has to be marked.
If Euro pallets are used, the setup would be a run-up of 15 pallets stacked 5 long and 3 high and a landing platform of 9 pallets stacked 3 long and 3 high.
Thus a minimum of 25 Euro pallets would be needed.
A Long Jump on Platform competition needs a minimum area of 20x2 meters.

\subsection{Judging}
The rider must start on the run-up and land on the landing platform without touching the ground.
After landing, the rider must remain mounted and on the landing platform or in the 3 meter half circle for 3 seconds, as counted by a judge.
The rider may do any form of idling, hopping or stillstanding during the 3 seconds.
Once the judge has counted 3 seconds, the jump is complete.
There has to be at least one judge.
