\chapter{Event Organizer Rules}

\section{Venue}
A smooth area of sufficient size must be set aside to run Jump events.
This must include a 15 m runway and a 3 m landing area for for both High Jump events, a 25 m runway and an 8 m landing area for Long Jump on Track.
The Event Host must organize the competition where marking tires and metal pedals are allowed.
If the venues are outdoors, plans must be made to deal with inclement weather.
Using an indoor venue can eliminate this problem.
The track must be available for enough days to allow for inclement weather.

\section{Officials}

The host must designate the following officials for Jump Events:
\begin{itemize}
\item Jump Director
\item Chief Judge
\item Judge
\end{itemize}

\begin{comment2016}
\section{Communication}
\end{comment2016}


\begin{comment2016}
\section{Age Groups}

Should this inherit the track age groups?
\end{comment2016}

\section{Practice}

The organizer should provide a place and equipment similar to those used for the official competition so the riders can practice before making their official attempts.
The equipment should be available during the whole length of the event, and even before if the organizer decides so.

For bigger events such as Unicon, national or continental events, the organizer must provide said equipment.

\section{Setup: High Jump over Bar}
Around the High Jump over Bar apparatus a circle with a radius of 3 meters must be marked.
This circle is start and finish line.
The bar must be held loosely in the jumping apparatus so it can fall or break away if the rider does not complete the desired height.
Magnetic systems are not allowed.
The bar shall have a minimum diameter of 2cm and a maximum bending of 2 cm.
(The bar may sag no more than 20 mm at its lowest point.)
The bar must be sufficiently long such that minimum distance between the two apparatuses holding the bar is 2 meters.

\section{Setup: High Jump onto Platform}
The structure consists of two parts: a platform and a landing surface.
The top surface of the platform must be at least 120 x 80 cm in size.
The sides of the platform must be nearly perpendicular with the ground to ensure its presence does not hamper riders.
The landing surface consists of a flat piece of wood firmly affixed to the top of the platform that is 120 x 80 cm in size.
If the top surface of the platform is larger than 120 x 80 cm, the landing surface must be attached to the platform such that one of the 120 cm sides is aligned with the edge of the platform.
The method of attachment must be chosen so that it does not interfere with riders during jump attempts.
The structure should be sturdily built and shimmed so that there is minimal motion when jumps are attempted.
The front of the platform must be covered by a wooden plate that extends a minimum of 60cm from the top of the platform down.
This cover must be used when riders are jumping 60cm or higher.

A 3 meter safety ring must be marked out around all sides of the platform where no persons may enter during a jump attempt.
Organizers may choose to mark out a 4 m ring where only select persons may enter during a jump attempt (e.g.\ photographers or judges).
A 15 m runway must also be cleared of persons for riders that will roll into the jump.

\section{Setup: Long Jump on Track}
The riding area consists of a start line, a jump marker, a landing marker and a finish line beyond the jump marker.

The finishing line should be at least 4 meters from the landing marker but no more than 8 meters away.
We suggest that judges set up the finishing line 8 meters from the jump marker and move it further away if need during longer jumps.
Riders must ride or hop across the finish line for the attempt to count.
Successfully crossing the finish line is judged the same as in racing (see section \ref{sec:track_finishes}).
The start line must be a minimum of 25 meters in front of the jump marker to allow the riders to accelerate.
There must be an area behind the finishing line which is a minimum of 7 meters long and 2 meters wide as safety zone.
Riders may use all or part of the 25 meters between start line and jump marker.
Riders are also allowed to start from beside to be able to do accelerated side jumps.
Markers for takeoff and landing (jump marker and landing marker) must consist of a material which cannot be deformed in order to have the same conditions for all riders.
The markers must be at least 1.20 meter in width (across the runway), no more than 10 mm in height (above the runway), and no less than 5 centimeters in depth (front to back).
A Long Jump on Track competition needs a minimum area of 40x2.5 meters.

\section{Setup: Long Jump on Platform}
The jump area consists of a takeoff platform and landing platform of the same height, width and length.

Both platforms should be fixed together to prevent movement throughout the event.
EUR pallets (120cm x 80cm x 14.5cm) are recommended and should be covered with plywood or a similar material of sufficient friction to allow for safe landings.
If EUR pallets are used the event requires 6 pallets however spares are recommended.

\textbf{Take-off and Landing platforms:}\\
The dimension of the takeoff and landing surface should be 120 cm x 80 cm (EUR pallet) with the platforms orientated parallel along the 120 cm side to ensure that riders are jumping from or landing onto the 120 cm side.
The platforms must be of equal height between 36cm and 45cm (3 pallets) high.
If EUR pallets are not available the platform should be marked to ensure the surface dimensions that may be used is no larger than 120cm x 80cm.
The riding area consists of a run-up platform and landing platform of the same height and width.
