\chapter{Event Organizer Rules}

All Track Other events follow the Track Racing event organizer rules, unless otherwise noted.

\section{Communication}

If a large convention advertises events with the names of the ones detailed in this section, they must use the rules provided here.
If hosts desire to do variations on these rules, the events must be labeled accordingly.
Example: ``Track Gliding; Modified''.
In cases such as this, hosts must remember to provide detailed rules for these events at the same time the events are announced.

\section{Age Groups}

 %comment2016 Do we need this?
Age group breakdown is also up to the host.

\section{Accuracy of Results}

For all Gliding and Coasting disciplines where the distance is measured, unless the distance is an exact 0.1 meter, the distance shall be converted and recorded to the next shorter 0.1 meter, e.g.\ 34.56 m shall be recorded as 34.5 m.
When two riders reach the same distance, it shall be determined to be a tie and the tie shall remain and gets published as such.

For all Slow disciplines and Stillstand, unless the time is an exact 0.1 second, the time shall be converted and recorded to the next shorter 0.1 second.
When two riders reach the same time, it shall be determined to be a tie and the tie shall remain and gets published as such.

For all other events mentioned in this chapter, unless the time is an exact 0.01 second, the time shall be converted and recorded to the next longer 0.01 second.
In the event that there is a tie where an award and/or a place in the final is at stake, if a photo finish system was used, the image of this system shall be used to decide on the placings.
In this case, the note (Photo Finish: +0.00X) is printed on the results list next to the official time. In other cases it shall be determined to be a tie and the tie shall remain and gets published as such.

\textbf{Example:} If two riders have reached a time of 0:07.08 and the image of the photo finish system shows a difference of 0.006 seconds, the following will be printed on the result list:\\
\begin{tabular}{l l l}
1st Place & Rider 1 & 0:07.08 \\
2nd Place & Rider 2 & 0:07.08 (Photo finish: +0.006)\\
\end{tabular}
