\chapter{Event Organizer Rules}

\section{Venue}
Very dangerous sections should be secured (for example by removing sharp stones/branches from areas where riders are likely to fall/run into due to the physics of the course).

Downhill and Cross Country courses must be rated in advance by two people using the appropriate IUF Muni Difficulty Scale (see the \href{https://unicycling.org/files/iuf-muni-difficulty-scale.pdf}{IUF publication} and \href{https://muni-scale.info/}{online calculator}).
Ratings and their underlying data must be published at least seven days prior to the event.

\subsection{Uphill}
Uphill courses must be primarily uphill but may include flat or downhill sections.
At Unicon, if the terrain is technically easy (i.e.\ smooth to ride, no or only very small rocks and roots), the main course has to be longer than 500 m or have a height difference greater than 100 m.
If the terrain is more difficult, the course may be shorter or have less elevation gain.
The event hosts may consider additionally offering Beginner/Advanced categories competing on shorter and easier tracks or on selected parts of the main course.
It is recommended that Uphill courses at other types of events or competitions adhere to these rules as well.

\subsection{Downhill}
Downhill courses must be primarily downhill but may include flat or uphill sections.
Recommended course length is 2.5 km, or 1 km at a minimum, depending on available terrain, trails and schedule time.
The overall course difficulty must be rated with the Unicycle Downhill Scale.
A minimum score of 20 points is recommended.
Courses with scores below 15 points should be labeled ``beginner downhill'' for clarity.

\subsection{Cross Country}
A Cross Country course should be held primarily off-road but may consist of gravel and paved roads.
There is freedom for the amount of downhill, flat and uphill sections, as long as the course is not primarily downhill or uphill.

\textbf{Special recommendations for Unicon and Continental Championships:}\\
There are four main categories based on distance and difficulty.
The overall course difficulty must be rated with the Unicycle XC Scale.
If multiple laps need to be completed, then the total distance is the basis for the rating.

Main categories:
\begin{enumerate}
\item \textbf{Beginners:} distance 3-7 km, max 15 points Unicycle XC Scale
\item \textbf{Intermediate:} distance 8-13 km, 16-21 points Unicycle XC Scale
\item \textbf{Elite:} distance 15-25 km, minimum 24 points Unicycle XC Scale, suggested to last 1h30min for fastest riders
\item \textbf{Marathon:} longer than 30 km, suggested to last 3-5 hours for fastest riders
\end{enumerate}

The elite category with subcategories Male and Female is required to be run.
Other categories are run on the host discretion.

The Elite Cross Country course should be designed such that the fastest rider completes the race in about 1 hour 30 minutes, but never less than 1 hour.
The course should be chosen such that geared riders do not have an undue advantage.
Unrideable sections should be avoided to prevent riders from walking however uphill sections may include technical difficulties.
Long and gentle uphills are preferred over short and steep ones.
If only shorter trails are available, riders can be required to complete multiple laps of the course.

\section{Officials}

The host must designate the following officials for each muni race:
\begin{itemize}
\item Muni Director
\item Referee
\item Starter
\item Course Marshall
\end{itemize}

\section{Communication}

\begin{comment2016}
Some thoughts on what might need to be communicated:
\begin{itemize}
\item false start method
\item age groups
\item starting method
\item timing method?
\item course map?
\item results
\end{itemize}
\end{comment2016}

The host must publish two lists of results for each discipline after the competition: Age group based ranking and overall ranking (separating male/female).

If the hosts wish to include events other than the first three (Up, DH, XC), they must remember to provide detailed rules for these events at the same time the events are announced.

Details of all muni events must be published as soon as they are known.
This is to provide competitors with the information they need to train, and to help them prepare the appropriate unicycles.
These are major needs for attendees from far away.
Necessary details depend on the event, but include things like course length, elevation and elevation change, steepness, level of terrain difficulty, amount of turns, riding surfaces, course width, etc.
Course details should be published as soon as possible and a GPS-track must be published on the web site at least one month prior to the convention.
Ratings of courses (using the IUF Muni Difficulty Scale) and their underlying data must be published at least seven days prior to the event.
It is acceptable to publish tentative courses while waiting for permits to be approved, etc.

\section{Age Groups}

\begin{comment2016}%comment2016 these age group rules are awful
Is this really meant to be ``maximum''?
\end{comment2016}

Age groups must be offered as male and female age group.
There must not be any age group specific restrictions on equipment.
The following age groups are the maximum allowable for muni competitions:

\begin{tabular}{|l|l|}
\hline
Under 15 & Youth \\
\hline
15-16 & Juniors \\
\hline
17-18 & Rookies \\
\hline
19-29 & Elites \\
\hline
30-49 & Masters \\
\hline
50+ & Veterans \\
\hline
\end{tabular}

\section{Route Signaling}
Courses must be clearly marked.
At each intersection, the correct direction must be indicated by at least one of the following options:
\begin{itemize}
\item \textbf{Option 1:} Painting or chalk marking (only if authorized by authorities).\\
Arrows showing the correct course direction must be drawn 5 to 10 meters before the intersection, at the intersection, and 5 to 10 meters after the intersection.
Crosses must be drawn on each wrong direction at the intersection and 5 to 10 meters after the intersection on the wrong paths.
\item \textbf{Option 2:} Using bands (also known as barrier tape).\\
Small pieces of bands (shorter than 1 m) are used instead of arrows to show the right way to go.
Longer pieces of bands barring the wrong paths over their entire width are used instead of crosses.
These bands can lie on the ground if they cannot be hung in the air because of any restriction.
\item \textbf{Option 3:} Using any other clear signaling method.\\
Chipped wood or large signs with printed arrows or crosses are examples of other acceptable method.
Similar to options 1 and 2, signaling must be placed 5 to 10 meters before, at, and 5 to 10 meters after any intersection, as well as signaling very clearly any wrong direction at the intersection, and 5 to 10 meters after it on the wrong paths.
\end{itemize}
Any element of the route such as the ground, trees, rocks, or barriers can be used as marking or for hanging bands, as long as the result is easily visible and not likely to be erased/removed by the passage of riders, other occupants, or weather.
If the weather forecast predicts rain, option 2 or 3 is preferred over option 1.
At major intersections, having a volunteer signaling the correct way, in addition to marks or bands, is highly recommended.
Any signaling option can also be used on any long section between intersections, in order to confirm to riders that they are still on the right track.

If authorized by authorities, an effort must be made to mark the courses a few days before they occur, so riders can practice on it.

Any change in the track from the one given in the website/program book must be announced to the riders by email as soon as this is known, even if this is a short amount of time before the race.
If the change occurs the day of the race, riders must also be told on the start line that there has been a change.

\section{Practice}

For all muni races, every rider must get the chance of at least one test run to get familiar with the track before the actual race.
If possible, the track should be open for training during all days of the event prior to the race.
For multi-day events the muni competitions should take place during the second half of the event in order to give riders more time to practice on the course.
The race cannot be held without the possibility of practice.

\section{Race Configuration}

For uphill and downhill races riders should race one at a time, released at regular time intervals.
If the schedule has a small time window for the race, riders should be run in heat sizes that allow passing on the course, and do not bottleneck at the beginning.

For a downhill course length less than 2 km, two separate runs should be held.
In this case the ranking of the riders is based on the fastest of the two runs.

For the uphill race, either one or two runs can be held.
In the case of two runs, the ranking of the riders is based on the fastest of the two runs.

\section{Starting Configuration  \label{sec:muni_starting-configuration}}

There are three different types of starting modes, that can be used in muni races.
\begin{enumerate}
\item \textbf{Mass starts:}\\
All riders start at the same time.
Mass starts must not be used when the race duration is expected to be shorter than 30 minutes.
The track must provide sufficient space for passing in the first section, so that the field of starters is aligned before the track narrows down.
Space for passing must be given along the track.
Mass starts with more than 40 riders have to be split to avoid accidents.
\item \textbf{Heat starts:}\\
Groups of riders start at intervals that can vary from 30 seconds to a few minutes.
The maximum number of riders per heat is determined by the average width of the first 100 m of the track.
There can be one rider for each meter in width.
The first heats must be separated based on gender with the first heat consisting of the top males and the second heat consisting of the top females.
After the top males begin, there must be a minimum 10 minute time interval before the top females start.
After the top females start there must be a minimum 5 minute time interval before the next heat begins.
\item \textbf{Individual starts:}\\
Individual riders start at intervals that can vary from 30 seconds to a few minutes.
\end{enumerate}

\section{Starting Order}

The fastest riders should always start first, regardless of the starting mode.
The order can be determined by seeding runs or another method.

\section{False Starts}

There are several options on how to deal with false starts:
\begin{itemize}
\item \textbf{Time Penalty:}\\
In case of a false start, the heat is not restarted.
If a false start occurs by one or multiple riders, these riders receive a time penalty (10 seconds is recommended).
\item \textbf{One False Start Allowed Per Rider:}\\
In case of a false start, the heat is restarted.
Any rider(s) who caused their personal first false start may start again.
Any rider(s) causing their personal second false start are disqualified.
\item \textbf{One False Start Allowed Per Heat:}\\
In case of a false start, the heat is restarted.
For the first false start of a particular heat, all riders may start again.
Thereafter, any rider(s) causing a false start are disqualified.
\end{itemize}

It is highly recommended to use the time penalty method for all different kinds of starts (i.e.\ mass, heat, and individual starts as described in section \ref{sec:muni_starting-configuration}).
If no false start rules are announced prior to the race, this is the default method.
In case of a purposeful false start several seconds before the official start, violating rider(s) should be disqualified at the discretion of the Starter.

At Unicons and continental championships, a camera has to be set up to record the start of all waves from a suitable angle.
The official footage from this camera serves as evidence in case of disputed false start events or other start-line related protests after the race.

Restarting heats (options 2 \& 3) should be avoided.
If a heat has to be restarted for any reason, the Starter will immediately recall the riders, for example by firing a gun or blowing a whistle or any other clear and pre-defined signal.

\section{Finish Camera}
At Unicons and Continental championships, a camera has to be set up to record the finish line from a suitable angle.
The official footage from this camera serves as evidence in case of dismounts over the finish line or other finish-line protests after the race.
