\chapter{Overview \label{chap:muni_racing}}

\section{Definition}

\oldrule{4.1}
Mountain unicycling (muni) refers to off-road races over any type of terrain.
Terrain can be anything from dirt to paved areas, hills, ditches, curbs, rocks, sand, mud, or grass (Mountains are not required).

\oldrule{4.2}
Mountain unicycling competitions usually involve Uphill (section \ref{sec:muni_uphill}), Downhill (section \ref{sec:muni_downhill}), and Cross Country (section \ref{sec:muni_xc}), but other events such as North Shore (a technical competition) and Slopestyle (a trick/style competition) can be included.

\oldrule{4.1}
\textit{For purposes of these rules, mountain unicycling (muni) refers to off road races over any type of terrain.
Races can vary from a single heat race with all riders starting together, to a time trial type of arrangement with riders going individually, at intervals.
Mountains are not required.
Terrain can be anything from dirt to paved areas, hills, ditches, curbs, rocks, sand, mud, or grass.
Unless otherwise noted, there are no restrictions on wheel size, crank arm length, brakes or gearing.}

\section{Rider Summary}

\textbf{This section is intended as an overview of the rules, but does not
substitute for the actual rules.
\begin{itemize}
\item You must wear shoes, knee pads, gloves, and helmet.
\item Personal music systems are not allowed.
\item Water and food stations are at the discretion of the host.
\item Muni racing events have no wheel size, crank length, or gearing 
requirements.
\item You must be on time for the start of your race.
\item Be aware of the rules regarding passing, dismounts, and illegal riding.
\end{itemize}}