\chapter{Judges and Officials Rules}

\begin{framed}
We think that the officials for each event need to be specified in event chapter, not in chapter 1. For this event, we have taken the approach that the rules will specify the scope of responsibility for the official.
\end{framed}

\section{Muni Director}

\begin{framed}
This and the next section were based on text from Chapter 1, reworded to make it specific to muni.
\end{framed}

\textbf{The Muni Director is the head organizer and administrator of muni
events.  With the Convention Host, the Muni Director determines the course,
obtains permis, interfaces with the community, and determines the system
used to run the event.  The Muni Director is responsible for the logistics and
equipment for all muni events.  With the Referee, the
Muni Director is in charge of keeping events running on schedule, and
answers all questions not pertaining to rules and judging.  The Muni
Director is the highest authority on everything to do with the muni
events, except for decisions on rules and results.}

\section{Referee}

\textbf{The Referee is the head muni official, whose primary job is to make sure
the competitors follow the rules.  The Referee makes all final decisions
regarding rule infractions. The Referee is responsible for resolving
protests.}

\section{Dismounted Riders}
\oldrule{5.10.1}
The following penalties apply if riders disregard \textbf{the rules about dismounts}:
\begin{itemize}
\item Riders get an immediate time penalty of five seconds when they intentionally run or walk fast, not recovering from a fall.
A judge must clearly indicate when the time penalty starts and when the rider may continue, for example by blowing a whistle and counting down from five.
\item Riders get disqualified immediately when they do not stop and wait five seconds after the judge's indication.
The disqualification should be signaled to the rider immediately by a judge, for example by blowing a whistle twice.
\item Judges must be trained and tested to correctly enforce these rules.
Riders must be informed about the type of signaling prior to the race.
\end{itemize}
