\chapter{General Rules and Definitions}
\section{General}
This Rulebook is intended to govern all unicycle competition sanctioned by the International Unicycling Federation, and can be used as a guideline for other competitions.
The major sections are General, Racing, Artistic, Games, Trials and Skill Levels.
Various charts and forms that supplement these rules may be published separately.

\subsection{These Are Official IUF Rules}
All IUF Unicons (International Unicycling Conventions) must abide exclusively by these rules.
Further rules may be added to cover specific situations, but they may not override the IUF rules without prior approval by the IUF Board of Directors.
All additional rules must be published well in advance of international competition, and published together with registration forms.

\subsubsection{Updating This Rulebook}
This publication should be updated after every Unicon.
The IUF Rulebook Chair will head the committee, but may optionally name a sub-committee.
The Rulebook Committee will officially start meeting at the close of the Unicon, though the Chairperson can open it before, to take advantage of having so many persons physically together.
The Committee should finish their business and make their specific proposals 

\subsection{Convention Aspect of Unicon}
All competitions at a Unicon need to make every effort to have equal time for the convention side of Unicon by involving
as many competitors as possible and making the event spectator-friendly for other Unicon participants as well as non-
unicyclists. Any of the following are examples to achieve this goal:
\begin{itemize}
  \item{Workshops related to the event}
  \item{Fun competitions based on the event}
  \item{Instant results for the spectators}
  \item{Ways for other competitors to be introduced to the event}
  \item{Entertainment during breaks in the competition (such as half time entertainment)}
  \item{Schedule of the events posted in multiple places}
\begin{itemize}

\subsection{Host’s Option - Unicon}
Unicon should include at least one event from each of the following event groups. 
Hosts are free to add events, age groups or variations that do not appear here, as long as there is no conflict with the existing rules. 
When in doubt contact the IUF Rules Committee.

