\chapter{Judges and Officials Rules}

\oldrule{2.3}
The Referee has final say on whether a rider’s safety equipment is sufficient. The Starter will remove from the starting line-up any riders not properly equipped to race, including riders with dangerously loose shoelaces. 

\section{Racing Officials}

\begin{framed}
All officials wording needs to be reviewed for consistency.  Some sections
include wording about training, but others to no not.
\end{framed}

\oldrule{1.28}
\textit{\textbf{Race Director:} The Race Director is in charge of seeing that all equipment, forms, people, sound systems, and other requirements are taken care of before the convention starts.
Ideally, the Race Director is a member of the host organization, or has convenient access to the convention's locations.
The Race Director is responsible for the logistics, equipment, and scheduling for all racing events unless otherwise noted.}

\textbf{The Track Director is the head organizer and administrator of track events.  The Track Director is responsible for the logistics and
equipment for all track events.  With the Referee, the
Track Director is in charge of keeping events running on schedule, and
answers all questions not pertaining to rules and judging.  The Track
Director is the highest authority on everything to do with the track
events, except for decisions on rules and results.}

\textit{\textbf{Referee:} The Head racing official.
Makes all final decisions regarding race competitions.
Handles protests.
Makes sure racing areas and officials are trained and ready.
Works within the system set up by the Race Director for running the events.
The Referee should be someone very experienced in all aspects of unicycle racing, and must above all be objective and favor neither local, nor outside riders.
There can be separate Referees for different venues, or different categories of racing.}

\textbf{The Referee is the head track official, whose primary job is to make sure
the competitors follow the rules.  The Referee makes all final decisions
regarding rule infractions. The Referee is responsible for resolving
protests. The Referee makes sure other track officials are trained and ready.}

\textbf{Starter:} Starts races; explains race rules; calls riders back in the event of false starts.
Also checks riders for correct unicycles and safety equipment.

\textbf{Finish Line Judge: The finish line judge determines whether rides cross the finish line properly, according to the rules.}

\begin{framed}
We suggest removing the auxilliary personnel, because they refer to tasks, 
not actual officials.  F
\end{framed}

\textit{\textbf{Clerk:} Sets up riders in lanes before races.
Checks riders for correct unicycles and safety equipment.}

\textit{\textbf{Picker:} Assists Timers by observing riders' finishing order.
Watches for finish line dismounts.|

\textit{\textbf{Timer:} Takes the time of riders at the finish line.
Also watches for finish line dismounts.
Two timers may be used for 1st place, the average time being official.}

\textit{\textbf{Recorder:} Writes down place and time of each rider after each event.
Riders must not leave the finish area until the Recorder has gotten their numbers and information.}

\textit{\textbf{Runner:} A general helper who brings racing forms from start to finish line, to tabulators, and to announcer.}

\textit{\textbf{Tabulator:} Processes all race results; totals points; prepares awards for presentation.}

\textit{\textbf{Results Poster:} Puts up Race Recording and Results Sheets for all to see, and marks the time.}

\textit{\textbf{Announcer:} Operates public address system, announces race results, and calls riders for upcoming races.}

\textit{\textbf{IUF Slalom and Slow Race Operators:} Run and administrate these two events in an area separate from the track, doing most of the above jobs for each.}

\section{Training Officials}
\oldrule{1.20}
As the rules state, competitions cannot be started until all key track officials have been trained and understand their tasks.
For Racing, the Referee is in charge of making sure this happens.

\section{Starter Responsibilities\label{sec:track_starter}}

\oldrule{2.4}
There should be about 3/4 second between each element in the count, with the same amount of time between each of them.
Starters should practice this before the races begin.
Timing of the count is very important for an accurate start.
This count can be in the local language, or a language agreed upon before competition starts.

Riders start mounted, holding onto a starting post or other support.
Unicycle riders need to be leaning forward before the starting gun fires, so the Starter will give a four-count start.
Example: ``One, two, three, BANG!''
This allows riders to predict the timing of the gun, for a fair start.

As an alternative a start-beep apparatus can be used.
In that case we have a six-count start.
Example: ``beep - beep -beep - beep - beep - buup!''
The timing between beeps is one second.
The first 5 beeps have all the same frequency.
The final tone (buup) has a slightly higher frequency, so that the racer can easily distinguish this tone from the rest.

\oldrule{2.5}
If a heat has to be restarted, the Starter will immediately recall the riders, for example by firing a gun or blowing a whistle or other clear and predefined signal.
It is only the earliest false starting rider who gets assigned this false start and might get disqualified.

\oldrule{2.5}
There are two options on how to deal with false starts:
\begin{itemize}
\item \textbf{One False Start Allowed Per Rider:}
In case of a false start, the heat is restarted.
Any rider(s) who caused their personal first false start may start again.
Any rider(s) causing their personal second false start are disqualified.
\item \textbf{One False Start Allowed Per Heat:} 
In case of a false start, the heat is restarted.
For the first false start of a particular heat, all riders may start again.
Thereafter, any rider(s) causing a false start are disqualified.
This option should not be used without an electronic false start monitoring system.
\end{itemize}

\section{Finish Line Judge Responsibilities}

\subsection{Judging Finish Line Dismounts}
\oldrule{2.6.1}
One or more officials are required at the finish line to judge dismounts in all races where dismounting is allowed.
These officials must be appointed by the racing referee so they fully understand their crucial job.
The finish line judges are the voice of authority on whether riders must remount and cross the finish line again.
Any riders affected must be clearly and immediately signaled to return to a spot before the finish line, remount without overlapping the finish line, then ride across it again.
The path for backing up may involve going around any finish line timing or optical equipment to prevent data problems for other riders in the race.

\subsection{Timing Penalty For Finish Line Dismounts}
\oldrule{2.6.2}
In electronically timed races, it's possible that no time will be recorded for the rider's successful finish.
Instead of recording an actual finish time, the rider's time will be recorded as 0.01 seconds faster than the next rider to cross the line after their remount and crossing.
If the rider in question is the last one on the track, the time recorded should be their actual time crossing the finish line after their remount.
