\chapter{Overview}

\section{Definition}

\textbf{Street is a skills competition where riders perform 
tricks on urban obstacles (stair sets, platforms, ramps, rails, etc.)
Riders are judged on the skill and creativity of moves and combinations they do.
The competition consists of preliminary rounds, followed by final.}

\oldrule{6.1}
\textit{
Street is about using the environment, such as ramps, rails, stairs and platforms, to do tricks in the style of skateboarding and.
Riders are judged on the skill and creativity of moves and combinations they do.}

\oldrule{6.24}
\textit{Street unicycling involves riders performing tricks using obstacles including stairs, boxes, and rails.
There is an emphasis on using the environment creatively and completing tricks smoothly and consistently.}

\section{Rider Summary}

\textbf{This section is intended as an overview of the rules, but does not
substitute for the actual rules.
\begin{itemize}
\item Be aware of the deadline for entry.
\item You must be on time for your scheduled performance.
\item You must wear shoes, shin pads, and helmet.
\item You may bring your own music, but music is not judged.
\item The competition is divided into zones, with different obstacles in each zone.
\item Riders compete in a group of three or four, and take turns riding the obstacles.
\item A rider may skip their turn, due to injury or equipment problem.
\item You should be aware of the rules regarding equipment failure.
\item Judges rank the riders according to the number of tricks done and the difficulty.
\item The top five or six riders are chosen to participate in the final.
\end{itemize}}